Naturalness and Long-lived Beyond the Standard Model Particles

### 2017-07-28 tex/draft.tex
Edit SUSY section

\section{Supersymmetry}
    Supersymmetry (SUSY) introduces a new symmetry in which every SM field fits into a larger multiplet with an inherent symmetry between bosons and fermions. In its simplest form, SUSY results in one new boson for every SM fermion and one new fermion for every SM boson \fxnote{and an extra Higgs field - probably should mention it. Make sure I understand it either way}. The increase in particle multiplicity calls for a new naming convention: the spin-0 superpartners to the SM fermions are referred to as sfermions (e.g. stop or smuon), and the spin-$\frac{1}{2}$ superpartners to the SM bosons are given the name of their SM counterpart with "ino" tacked on to the end (e.g. Higgsino or wino). Together the superpartners to the SM particles are known as sparticles.

    When calculating contributions from loop diagrams, one finds that fermion loops differ in sign from boson loops, which means that in SUSY every bosonic correction to the Higgs mass is cancelled by a fermionic correction and vice versa. In fact, exact supersymmetry predicts that the bare Higgs mass is uncorrected to all orders in perturbation theory. \fxnote{show example diagram with contribution?}

    But the lack of sparticles in collider searches and everyday life tells us that SUSY must be broken. If SUSY were exact, every superpartner would have the exact same mass as its SM counterpart, which means all sparticles would have been found at the same energy scales as SM particles. The mechanism and scale of SUSY breaking is an open question with many competing answers that achieve different theoretical goals and lead to different TeV-scale phenemena. A handful of models with particular approaches to SUSY breaking that are particularly prominant or particularly relevant to naturalness or long-lived particles are gauge-mediated susy breaking (GMSB), anomaly-mediated SUSY breaking (AMSB), Split SUSY, and R-parity violating SUSY (RPV). The details of these models will be explored as needed throughout this paper.

>>>>

\section{Supersymmetry}
    \fxnote{seems like I should mention that naturalness favors lighter sparticles}  
    Supersymmetry (SUSY) introduces a new symmetry in which every SM field fits into a larger multiplet with an inherent symmetry between bosons and fermions\fxnote{I need to be better at explaining exactly what this means}. In its simplest form, SUSY results in one new boson for every SM fermion, one new fermion for every SM boson, and one new Higgs doublet.\fxnote{make sure I understand the new Higgs doublet} The increase in particle multiplicity calls for a new naming convention: the spin-0 superpartners to the SM fermions are referred to as sfermions (e.g. stop or smuon), and the spin-$\frac{1}{2}$ superpartners to the SM bosons are given the name of their SM counterpart with "ino" tacked on to the end (e.g. Higgsino or wino). Collectively the superpartners to the SM particles are known as sparticles.\fxnote{cite martin}

    When calculating contributions from loop diagrams, one finds that fermion loops differ in sign from boson loops, which means that in SUSY every bosonic correction to the Higgs mass is cancelled by a fermionic correction and vice versa. In fact, exact supersymmetry predicts that the bare Higgs mass is uncorrected to all orders in perturbation theory, as shown below.

    \fxnote{exact SUSY diagrams and correction term goes here}

    Exact SUSY also requires that sparticles have the same mass as their SM counterparts, so the lack of sparticles in collider searches implies that SUSY must be broken. With broken susy, the same diagram as above produces a correction of the form

    \fxnote{broken SUSY correction term goes here}

    The mechanism and scale of SUSY breaking is an open question with many competing answers that achieve different theoretical goals and lead to different TeV-scale phenemena. A handful of models with particular approaches to SUSY breaking that are particularly prominant or particularly relevant to naturalness or long-lived particles are gauge-mediated susy breaking (GMSB), anomaly-mediated SUSY breaking (AMSB), Split SUSY, and R-parity violating SUSY (RPV). The details of these models will be explored as needed throughout this paper.
